Rethinking how students learn to design programs

### 2016-04-23 manuscript-body.tex
Finish citations in Introduction

But, for all the data these methods collect their focus is primarily
what Jadud (2006) calls students' ``syntactic'' struggles: the challenge
of articulating well-formed statements a compiler can properly parse.
This focus on syntax-level struggles leads to both a methodological and
theoretical trade-off: we can see in detail how students struggle with
wording, but we see less of how they struggle with the meaning and
intent of their code. Syntactical analyses necessarily ignore the
specific \emph{content} of students' code because they abstract
meta-information about the event: error type, error location, frequency
of error.

As an analogy, suppose that instead of studying computer science
students we were studying screenwriting students. Further suppose we
have a system to track data whenever students save their screenplays and
scripts. For each save, we get a copy of the entire script at that time.
We can also run an automated analysis to check whether they've properly
formatted slug lines\footnote{In a screenplay, a slug line gives
  information about where and when a scene takes place. For example:
  ``INT. FORTRESS OF SOLITUDE -- DAY.''}, put character names in
all-capital letters, properly numbered scenes, etc. We might even know
where students are when they write (coffee shop, library, home, etc.).
If we mine that data, we could learn something about students'
\emph{screenwriting behavior}, but only at the level of how they
struggle with screenwriting syntax. Using only error-based data it's
much harder to answer questions such as:

\begin{itemize}
\item
  How does this student construct a scene or handle dialog in their
  script?
\item
  Does the student seem to have a grasp of pacing, story beats, and
  efficient exposition?
\item
  How does their writing manage and develop character arcs?
\end{itemize}

@@

\begin{itemize}
\item
  Do they use particular techniques to ``break story'' and decompose a
  narrative into its key beats (index cards? Whiteboard?)
\item
  Do they participate in a writing group?
\item
  How do they respond to people giving them notes on revising a script?
\end{itemize}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Snapshotting students' code represents a cutting-edge way to resolve
  the way code---as a design artifact---evolves over time, \emph{but}
\item
  Code snapshots, as they're currently used, explore neither the
  totality of a student's design nor the rich context of that design's
  production, \emph{and}
\item
  There is a lack of parity between studies of how professionals design
  software and studies of how students do so,
\end{enumerate}

>>>>

But, for all the data these methods collect their focus is primarily
what Jadud \cite{jadud_methods_2006} calls students' ``syntactic'' struggles: the challenge
of articulating well-formed statements a compiler can properly parse.
This focus on syntax-level struggles leads to both a methodological and
theoretical trade-off: we can see in detail how students struggle with
wording, but we see less of how they struggle with the meaning and
intent of their code. Syntactical analyses necessarily ignore the
specific \emph{content} of students' code because they abstract
meta-information about the event: error type, error location, frequency
of error.

As an analogy, suppose that instead of studying computer science
students we were studying screenwriting students. Further suppose we
have a system to track data whenever students save their screenplays and
scripts. For each save, we get a copy of the entire script at that time.
We can also run an automated analysis to check whether they've properly
formatted slug lines\footnote{In a screenplay, a slug line gives
  information about where and when a scene takes place. For example:
  \texttt{``INT. FORTRESS OF SOLITUDE -- DAY.''}}, put character names in
all-capital letters, properly numbered scenes, etc. We might even know
where students are when they write (coffee shop, library, home, etc.).
If we mine that data, we could learn something about students'
\emph{screenwriting behavior}, but only at the level of how they
struggle with screenwriting syntax. Using only error-based data it's
much harder to answer questions such as:

\begin{itemize}
  \tightlist
\item
  How does this student construct a scene or handle dialog in their
  script?
\item
  Does the student seem to have a grasp of pacing, story beats, and
  efficient exposition?
\item
  How does their writing manage and develop character arcs?
\end{itemize}

@@

\begin{itemize}
  \tightlist
\item
  Do they use particular techniques to ``break story'' and decompose a
  narrative into its key beats (index cards? Whiteboard?)
\item
  Do they participate in a writing group?
\item
  How do they respond to people giving them notes on revising a script?
\end{itemize}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Snapshotting students' code represents a cutting-edge way to resolve
  the way code---as a design artifact---evolves over time, \emph{but}
\item
  Code snapshots, as they're currently used, explore neither the
  totality of a student's design nor the rich context of that design's
  production, \emph{and}
\item
  There is a lack of parity between studies of how professionals design
  software and studies of how students do so,
\end{enumerate}
